---
title: "Retail Strategy and Analytics-Task2"
author: "Rahul"
date: '2022-08-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(linewidth=80)
```


```{r knitr line wrap setup, include=FALSE}
library(knitr)
hook_output = knit_hooks$get("output")
knit_hooks$set(output = function(x, options)
{
 # this hook is used only when the linewidth option is not NULL
 if (!is.null(n <- options$linewidth))
 {
 x = knitr:::split_lines(x)
 # any lines wider than n should be wrapped
 if (any(nchar(x) > n))
 x = strwrap(x, width = n)
 x = paste(x, collapse = "\n")
 }
 hook_output(x, options)
})

```

```{r}
library(data.table)
library(ggplot2)
library(tidyr)
library(dplyr)
```
<style type="text/css">

body, td {
   font-size: 20px;
}
code.r{
  font-size: 16px;
}
pre {
  font-size: 16px
}
</style>

```{r}
setwd("E:/study/me/accenture")
data <- read.csv("QVI_data.csv")
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
```

## Select control stores

Calculate these measures over time for each store Add a new month ID column in the data with the format yyyymm.
```{r}
setDT(data)
data[, YEARMONTH := year(DATE)*100 + month(DATE)]
```


Next, we define the measure calculations to use during the analysis.For each store and month calculate total sales, number of customers, transactions per customer, chips per customer and the average price per unit.
```{r}
measureOverTime <- data[, .(totSales = sum(TOT_SALES),
                            nCustomers = uniqueN(LYLTY_CARD_NBR) ,
                            nTxnPerCust = uniqueN(TXN_ID)/uniqueN(LYLTY_CARD_NBR),
                            nChipsPerTxn = sum(PROD_QTY)/uniqueN(TXN_ID),
                            avgPricePerUnit = sum(TOT_SALES)/sum(PROD_QTY))
                        , by = c("STORE_NBR", "YEARMONTH")][order(STORE_NBR, YEARMONTH) ]
```
  

Filter to the pre-trial period and stores with full observation periods
```{r}
storesWithFullObs <- unique(measureOverTime[, .N, STORE_NBR][N == 12, STORE_NBR])
preTrialMeasures <- measureOverTime[YEARMONTH < 201902 & STORE_NBR %in%
storesWithFullObs, ]

```

Create a function to calculate correlation for a measure, looping through each control store.

```{r}
calculateCorrelation <- function(inputTable, metricCol, storeComparison) {
  calcCorrTable = data.table(Store1 = numeric(), Store2 = numeric(), corr_measure =
  numeric())
  storeNumbers <- unique(inputTable[, STORE_NBR])
  
  for (i in storeNumbers) {
    calculatedMeasure = data.table("Store1" = storeComparison,
                                   "Store2" = i,
                                   "corr_measure" = cor( inputTable[STORE_NBR == storeComparison,
                                                                     eval(metricCol)], inputTable[STORE_NBR == i,
                                                                                                eval(metricCol)]))
    calcCorrTable <- rbind(calcCorrTable, calculatedMeasure)
  }
  return(calcCorrTable)
}
```

Create function to calculate magnitude distance
```{r}
calculateMagnitudeDistance <- function(inputTable, metricCol, storeComparison)
{
  {
calcDistTable = data.table(Store1 = numeric(), Store2 = numeric(), YEARMONTH =
 numeric(), measure = numeric())
storeNumbers <- unique(inputTable[, STORE_NBR])
for (i in storeNumbers) {
calculatedMeasure = data.table("Store1" = storeComparison
, "Store2" = i, "YEARMONTH" = inputTable[STORE_NBR ==
 storeComparison, YEARMONTH]
, "measure" = abs(inputTable[STORE_NBR ==
 storeComparison, eval(metricCol)]
- inputTable[STORE_NBR == i,
 eval(metricCol)])
)
calcDistTable <- rbind(calcDistTable, calculatedMeasure)
}
}

## Standardise the magnitude distance so that the measure ranges from 0 to 1
  minMaxDist <- calcDistTable[, .(minDist = min(measure), maxDist = max(measure)),
by = c("Store1", "YEARMONTH")]
  distTable <- merge(calcDistTable, minMaxDist, by = c("Store1", "YEARMONTH"))
  distTable[, magnitudeMeasure := 1 - (measure - minDist)/(maxDist - minDist)]
  finalDistTable <- distTable[, .(mag_measure = mean(magnitudeMeasure)), by =
.(Store1, Store2)]
     return(finalDistTable)
}
```

Use the function you created to calculate correlations against store 77 using total sales and number of customers.
```{r}
trial_store <- 77
corr_nSales <- calculateCorrelation(preTrialMeasures, quote(totSales), trial_store)
corr_nSales[order(-corr_measure)]
```


```{r}
corr_nCustomers <- calculateCorrelation(preTrialMeasures, quote(nCustomers), trial_store)
corr_nCustomers[order(-corr_measure)]
```
Then, use the functions for calculating magnitude.
```{r}
magnitude_nSales <- calculateMagnitudeDistance(preTrialMeasures, quote(totSales),
trial_store)
magnitude_nCustomers <- calculateMagnitudeDistance(preTrialMeasures,
quote(nCustomers), trial_store)
```

Create a combined score composed of correlation and magnitude, by first merging the correlations table with the magnitude table
```{r}
corr_weight <- 0.5
score_nSales <- merge(corr_nSales, magnitude_nSales, by = 
          c("Store1","Store2"))[, scoreNSales := (corr_measure + mag_measure)/2 ]
score_nCustomers <- merge(corr_nCustomers, magnitude_nCustomers, by =
        c("Store1", "Store2"))[, scoreNCust := (corr_measure + mag_measure)/2]

score_nSales[order(-scoreNSales)]
```


```{r}
score_nCustomers[order(-scoreNCust)]
```

Combine scores across the drivers by first merging our sales scores and customer scores into a single table
```{r}
score_Control <- merge(score_nSales, score_nCustomers, by = c("Store1","Store2"))
score_Control[, finalControlScore := scoreNSales * 0.5 + scoreNCust * 0.5]
score_Control[order(-finalControlScore)]
```

Select control stores based on the highest matching store (closest to 1 but not the store itself, i.e. the second ranked highest store)
```{r}
control_store <- score_Control[Store1 == trial_store, ][order(-finalControlScore)][2, Store2]
control_store
```

Visual checks on trends based on the drivers
```{r}
measureOverTimeSales <- measureOverTime

pastSales <- measureOverTimeSales[, Store_type := ifelse(STORE_NBR == trial_store,
"Trial",
                                          ifelse(STORE_NBR == control_store,
"Control", "Other stores"))
                              ][, totSales := mean(totSales), by = c("YEARMONTH",
"Store_type")
                              ][, TransactionMonth := as.Date(paste(YEARMONTH %/%
100, YEARMONTH %% 100, 1, sep = "-"), "%Y-%m-%d")
                              ][YEARMONTH < 201903 , ]
ggplot(pastSales, aes(TransactionMonth, totSales, color = Store_type)) +
  geom_line() +
  labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")
```


Conduct visual checks on customer count trends by comparing the trial store to the control store and other stores.
```{r}
measureOverTimeCusts <- measureOverTime
pastCustomers <- measureOverTimeCusts[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
                                      ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, numberCustomers := mean(nCustomers), by = c("YEARMONTH", "Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/% 100, YEARMONTH %% 100, 1, sep = "-"), "%Y-%m-%d")
][YEARMONTH < 201903 , ]
ggplot(pastCustomers, aes(TransactionMonth, numberCustomers, color = Store_type)) +
  geom_line() +
  labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month")
```

## Assessment of trial

Scale pre-trial control sales to match pre-trial trial store sales
```{r}
scalingFactorForControlSales <- preTrialMeasures[STORE_NBR == trial_store &
YEARMONTH < 201902, sum(totSales)]/preTrialMeasures[STORE_NBR == control_store &
YEARMONTH < 201902, sum(totSales)]
## Apply the scaling factor
measureOverTimeSales <- measureOverTime
scaledControlSales <- measureOverTimeSales[STORE_NBR == control_store, ][ ,
controlSales := totSales * scalingFactorForControlSales]
```

Calculate the percentage difference between scaled control sales and trial sales
```{r}
percentageDiff <- merge(scaledControlSales[, c("YEARMONTH", "controlSales")],
                        measureOverTime[STORE_NBR == trial_store, c("totSales", "YEARMONTH")],
                        by = "YEARMONTH")[, percentageDiff := abs(controlSales-totSales)/controlSales]
```


As our null hypothesis is that the trial period is the same as the pre-trial period
```{r}
stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff])

## Note that there are 8 months in the pre-trial period hence 8 - 1 = 7 degrees of freedom
degreesOfFreedom <- 7

## We will test with a null hypothesis of there being 0 difference between trial and control stores.
percentageDiff[, tValue := (percentageDiff - 0)/stdDev
               ][, TransactionMonth := as.Date(paste(YEARMONTH %/% 100, YEARMONTH %% 100, 1, 
                                                     sep = "-"), "%Y-%m-%d")
][YEARMONTH < 201905 & YEARMONTH > 201901, .(TransactionMonth,tValue)]
```

Find the 95th percentile of the t distribution with the appropriate degrees of freedom to compare against
```{r}
qt(0.95, df = degreesOfFreedom)

measureOverTimeSales <- measureOverTime
```

Trial and control store total sales Create new variables Store_type, totSales and TransactionMonth in the data table.
```{r}
pastSales <- measureOverTimeSales[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, totSales := mean(totSales), by = c("YEARMONTH", "Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/% 100, YEARMONTH %% 100, 1, sep = "-"), "%Y-%m-%d")
][Store_type %in% c("Trial", "Control"), ]

```

Control store 95th percentile
```{r}
pastSales_Controls95 <- pastSales[Store_type == "Control",
 ][, totSales := totSales * (1 + stdDev * 2)
 ][, Store_type := "Control 95th % confidence
interval"]
```

Control store 5th percentile
```{r}
pastSales_Controls5 <- pastSales[Store_type == "Control",
 ][, totSales := totSales * (1 - stdDev * 2)
 ][, Store_type := "Control 5th % confidence
interval"]
trialAssessment <- rbind(pastSales, pastSales_Controls95, pastSales_Controls5)

```

Plotting these in graph
```{r}
ggplot(trialAssessment, aes(TransactionMonth, totSales, color = Store_type)) +
  geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,],
aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 , ymax =
Inf, color = NULL), show.legend = FALSE) +
  geom_line() +
  labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")
```

Compute a scaling factor to align control store customer counts to our trial store.Then, apply the scaling factor to control store customer counts.
```{r}
scalingFactorForControlCust <- preTrialMeasures[STORE_NBR == trial_store &
YEARMONTH < 201902, sum(nCustomers)] / preTrialMeasures[STORE_NBR ==
control_store & YEARMONTH < 201902, sum(nCustomers)]
measureOverTimeCusts <- measureOverTime
scaledControlCustomers <- measureOverTimeCusts[STORE_NBR == control_store,
][ , controlCustomers := nCustomers * scalingFactorForControlCust
][, Store_type := ifelse(STORE_NBR ==trial_store, "Trial",
ifelse(STORE_NBR == control_store,"Control", "Other stores"))]
percentageDiff <- merge(scaledControlCustomers[, c("YEARMONTH", "controlCustomers")],
measureOverTimeCusts[STORE_NBR == trial_store,c("nCustomers", "YEARMONTH")],
by = "YEARMONTH"
)[, percentageDiff := abs(controlCustomers-nCustomers)/controlCustomers]
```

As our null hypothesis is that the trial period is the same as the pre-trial period
```{r}
stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff])
degreesOfFreedom <- 7
```

Trial and control store number of customers
```{r}
pastCustomers <- measureOverTimeCusts[, nCusts := mean(nCustomers), by =
c("YEARMONTH", "Store_type")
                            ][Store_type %in% c("Trial", "Control"), ]

```

Control store 95th percentile
```{r}
pastCustomers_Controls95 <- pastCustomers[Store_type == "Control",
                                ][, nCusts := nCusts * (1 + stdDev * 2)
                                ][, Store_type := "Control 95th % confidence
interval"]
```

Control store 5th percentile
```{r}
pastCustomers_Controls5 <- pastCustomers[Store_type == "Control",
                                ][, nCusts := nCusts * (1 - stdDev * 2)
                                ][, Store_type := "Control 5th % confidence
interval"]
trialAssessment <- rbind(pastCustomers, pastCustomers_Controls95,
pastCustomers_Controls5)
```

Plot everything into one graph.
```{r}
ggplot(trialAssessment, aes(TransactionMonth, nCusts, color = Store_type)) + 
  geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,],
aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 ,
ymax = Inf, color = NULL), show.legend = FALSE) + 
  geom_line() + labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month")
```

## Trial store 86

Calculate the metrics below as we did for the first trial store.
```{r}
measureOverTime <- data[, .(totSales = sum(TOT_SALES),
                            nCustomers = uniqueN(LYLTY_CARD_NBR),
                            nTxnPerCust = (uniqueN(TXN_ID))/(uniqueN(LYLTY_CARD_NBR)),
                            nChipsPerTxn = (sum(PROD_QTY))/(uniqueN(TXN_ID)) , 
                            avgPricePerUnit = sum(TOT_SALES)/sum(PROD_QTY) ) , by = c("STORE_NBR", "YEARMONTH")][order(STORE_NBR, YEARMONTH)]

```

Using the functions we created earlier to calculate correlations and magnitude for each potential control store
```{r}
trial_store <- 86
corr_nSales <- calculateCorrelation(preTrialMeasures, quote(totSales),trial_store)
corr_nCustomers <- calculateCorrelation(preTrialMeasures, quote(nCustomers), trial_store)
magnitude_nSales <- calculateMagnitudeDistance(preTrialMeasures, quote(totSales), trial_store)
magnitude_nCustomers <- calculateMagnitudeDistance(preTrialMeasures, quote(nCustomers), trial_store)
```

Now, create a combined score composed of correlation and magnitude
```{r}
corr_weight <- 0.5
score_nSales <- merge(corr_nSales, magnitude_nSales, by = c("Store1", "Store2"))[ , scoreNSales := (corr_measure + mag_measure)/2]
score_nCustomers <- merge(corr_nCustomers, magnitude_nCustomers, by = c("Store1", "Store2"))[ , scoreNCust := (corr_measure + mag_measure)/2]

```

Finally, combine scores across the drivers using a simple average.
```{r}

score_Control <- merge(score_nSales, score_nCustomers, by = c("Store1","Store2"))
score_Control[, finalControlScore := scoreNSales * 0.5 + scoreNCust * 0.5]
```

Select control stores based on the highest matching store
```{r}
control_store <- score_Control[Store1 == trial_store,
][order(-finalControlScore)][2, Store2]
control_store
```

Conduct visual checks on trends based on the drivers
```{r}

measureOverTimeSales <- measureOverTime
pastSales <- measureOverTimeSales[, Store_type:= ifelse(STORE_NBR == trial_store, "Trial", ifelse(STORE_NBR== control_store, "Control", "Other stores"))][, totSales := mean(totSales), by = c("YEARMONTH", "Store_type")][, TransactionMonth:= as.Date(paste(YEARMONTH%/%100, YEARMONTH%% 100, 1, sep = "-"), "%Y-%m-%d")][YEARMONTH <210903] 

ggplot(pastSales, aes(TransactionMonth, totSales, color = Store_type)) +
  geom_line(aes(linetype = Store_type)) +
  labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")
```

Conduct visual checks on trends based on the drivers
```{r}
measureOverTimeCusts <- measureOverTime
pastCustomers <- measureOverTimeCusts[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, numberCustomers := mean(nCustomers), by = c("YEARMONTH", "Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/%
                                        100, YEARMONTH %% 100, 1, sep = "-"), "%Y-%m-%d")
][YEARMONTH < 201903 , ]

ggplot(pastCustomers, aes(TransactionMonth, numberCustomers, color = Store_type)) + 
  geom_line() + 
  labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month")
```

Scale pre-trial control sales to match pre-trial trial store sales
```{r}
scalingFactorForControlSales <- preTrialMeasures[STORE_NBR == trial_store &
YEARMONTH < 201902, sum(totSales)]/preTrialMeasures[STORE_NBR == control_store &
YEARMONTH < 201902, sum(totSales)]

## Apply the scaling factor
measureOverTimeSales <- measureOverTime
scaledControlSales <- measureOverTimeSales[STORE_NBR == control_store, ][ ,
controlSales := totSales * scalingFactorForControlSales]

## Calculate the percentage difference between scaled control sales and trial sales When calculating percentage difference
percentageDiff <- merge(scaledControlSales[, c("YEARMONTH", "controlSales")],
measureOverTime[STORE_NBR == trial_store, c("totSales", "YEARMONTH")],
by = "YEARMONTH"
)[, percentageDiff := abs(controlSales-totSales)/controlSales]

```

As our null hypothesis is that the trial period is the same as the pre-trial
```{r}
stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff])
degreesOfFreedom <- 7

```

 We only need data for the trial and control store.
```{r}
measureOverTimeSales <- measureOverTime
pastSales <- measureOverTimeSales[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, totSales := mean(totSales), by = c("YEARMONTH", "Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/%100, YEARMONTH %% 100, 1, sep = "-"), "%Y-%m-%d")
][Store_type %in% c("Trial", "Control"), ]

```

Calculate the 5th and 95th percentile for control store sales.Control store 95th percentile
```{r}
pastSales_Controls95 <- pastSales[Store_type == "Control",
][, totSales := totSales * (1 + stdDev * 2)
][, Store_type := "Control 95th % confidence interval"]

```

Control store 5th percentile
```{r}
pastSales_Controls5 <- pastSales[Store_type == "Control",
][, totSales := totSales * (1 - stdDev * 2)
][, Store_type := "Control 5th % confidence interval"]
```

create a combined table with columns from pastSales, pastSales_Controls95 and pastSales_Controls5
```{r}
trialAssessment <- rbind(pastSales, pastSales_Controls95, pastSales_Controls5)

## Plotting these in one nice graph
ggplot(trialAssessment, aes(TransactionMonth, totSales, color = Store_type)) +
  geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,],
aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 , ymax =
Inf, color = NULL), show.legend = FALSE) +
  geom_line(aes(linetype = Store_type)) +
  labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")
```

Scale pre-trial control customers to match pre-trial trial store customers
```{r}
scalingFactorForControlCust <- preTrialMeasures[STORE_NBR == trial_store &
YEARMONTH < 201902, sum(nCustomers)]/preTrialMeasures[STORE_NBR == control_store &
YEARMONTH < 201902, sum(nCustomers)]
## Apply the scaling factor
measureOverTimeCusts <- measureOverTime
scaledControlCustomers <- measureOverTimeCusts[STORE_NBR == control_store,
                                            ][ , controlCustomers := nCustomers
* scalingFactorForControlCust
                                            ][, Store_type := ifelse(STORE_NBR
== trial_store, "Trial",
                                      ifelse(STORE_NBR == control_store,
"Control", "Other stores"))
]
```

Calculate the percentage difference between scaled control sales and trial sales
```{r}
percentageDiff <- merge(scaledControlCustomers[, c("YEARMONTH",
"controlCustomers")],
                        measureOverTime[STORE_NBR == trial_store, c("nCustomers",
"YEARMONTH")],
                        by = "YEARMONTH"
                        )[, percentageDiff :=
abs(controlCustomers-nCustomers)/controlCustomers]
```

As our null hypothesis is that the trial period is the same as the pre-trial period
```{r}
stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff])
degreesOfFreedom <- 7
## Trial and control store number of customers
pastCustomers <- measureOverTimeCusts[, nCusts := mean(nCustomers), by =
c("YEARMONTH", "Store_type")
                              ][Store_type %in% c("Trial", "Control"), ]
## Control store 95th percentile
pastCustomers_Controls95 <- pastCustomers[Store_type == "Control",
                                ][, nCusts := nCusts * (1 + stdDev * 2)
                                ][, Store_type := "Control 95th % confidence
interval"]
```

Control store 5th percentile
```{r}
pastCustomers_Controls5 <- pastCustomers[Store_type == "Control",
                              ][, nCusts := nCusts * (1 - stdDev * 2)
                              ][, Store_type := "Control 5th % confidence
interval"]
trialAssessment <- rbind(pastCustomers, pastCustomers_Controls95,
pastCustomers_Controls5)
## Plotting these in one nice graph
ggplot(trialAssessment, aes(TransactionMonth, nCusts, color = Store_type)) +
  geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,],
aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 , ymax =
Inf, color = NULL), show.legend = FALSE) +
  geom_line() +
  labs(x = "Month of operation", y = "Total number of customers", title = "Total
number of customers by month")
```

## Trial store 88

Conduct the analysis on trial store 88.
```{r}
measureOverTime <- data[, .(totSales = sum(TOT_SALES),
nCustomers = uniqueN(LYLTY_CARD_NBR),
nTxnPerCust = uniqueN(TXN_ID)/uniqueN(LYLTY_CARD_NBR),
nChipsPerTxn = sum(PROD_QTY)/uniqueN(TXN_ID),
avgPricePerUnit = sum(TOT_SALES)/sum(PROD_QTY))
, by = c("STORE_NBR", "YEARMONTH")][order(STORE_NBR, YEARMONTH)]
```

Using the functions from earlier to calculate the correlation of the sales and number of customers of each potential control store to the trial store
```{r}
trial_store <- 88
corr_nSales <- calculateCorrelation(preTrialMeasures, quote(totSales),trial_store)
corr_nCustomers <- calculateCorrelation(preTrialMeasures, quote(nCustomers), trial_store)
## Use the functions from earlier to calculate the magnitude distance of the sales and number of customers of each potential control store to the trial store
magnitude_nSales <- calculateMagnitudeDistance(preTrialMeasures, quote(totSales), trial_store)
magnitude_nCustomers <- calculateMagnitudeDistance(preTrialMeasures, quote(nCustomers), trial_store)
## Create a combined score composed of correlation and magnitude by merging the correlations table and the magnitudes table, for each driver.
corr_weight <- 0.5
score_nSales <- merge(corr_nSales, magnitude_nSales, by = c("Store1", "Store2"))[ , scoreNSales := (corr_measure + mag_measure)/2]
score_nCustomers <- merge(corr_nCustomers, magnitude_nCustomers, by = c("Store1", "Store2"))[ , scoreNCust := (corr_measure + mag_measure)/2]
```

Combine scores across the drivers by merging sales scores and customer scores, and compute a final combined score.
```{r}
score_Control <- merge(score_nSales, score_nCustomers, by = c("Store1","Store2"))
score_Control[, finalControlScore := scoreNSales * 0.5 + scoreNCust * 0.5]
## Select control stores based on the highest matching store Select control store for trial store 88
control_store <- score_Control[Store1 == trial_store, ][order(-finalControlScore)][2, Store2]
control_store
```

For the period before the trial, create a graph with total sales of the trial store for each month, compared to the control store and other stores.
```{r}
measureOverTimeSales <- measureOverTime
pastSales <- measureOverTimeSales[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, totSales := mean(totSales), by = c("YEARMONTH","Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/% 100, YEARMONTH %% 100, 1, sep = "-"), "%Y-%m-%d")
][YEARMONTH < 201903 , ]
ggplot(pastSales, aes(TransactionMonth, totSales, color = Store_type)) +
geom_line(aes(linetype = Store_type)) +
labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")
```

For the period before the trial, create a graph with customer counts of the trial store for each month, compared to the control store and other stores.
```{r}
measureOverTimeCusts <- measureOverTime
pastCustomers <- measureOverTimeCusts[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, numberCustomers := mean(nCustomers), by = c("YEARMONTH", "Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/%
                                        100, YEARMONTH %% 100, 1, sep = "-"), "%Y-%m-%d")
][YEARMONTH < 201903 , ]
ggplot(pastCustomers, aes(TransactionMonth, numberCustomers, color = Store_type)) + 
  geom_line() + labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month")
```

Scale pre-trial control store sales to match pre-trial trial store sales
```{r}
scalingFactorForControlSales <- preTrialMeasures[STORE_NBR == trial_store &
YEARMONTH < 201902, sum(totSales)]/preTrialMeasures[STORE_NBR ==
control_store & YEARMONTH < 201902, sum(totSales)]

## Apply the scaling factor
measureOverTimeSales <- measureOverTime
scaledControlSales <- measureOverTimeSales[STORE_NBR == control_store, ][ ,controlSales := totSales * scalingFactorForControlSales]

## Calculate the absolute percentage difference between scaled control sales and trial sales
percentageDiff <- merge(scaledControlSales[, c("YEARMONTH", "controlSales")],measureOverTime[STORE_NBR == trial_store, c("totSales", "YEARMONTH")],by = "YEARMONTH")[, percentageDiff := abs(controlSales-totSales)/controlSales]

```

As our null hypothesis is that the trial period is the same as the pre-trial period
```{r}
stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff])
degreesOfFreedom <- 7
## Trial and control store total sales
measureOverTimeSales <- measureOverTime
pastSales <- measureOverTimeSales[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, totSales := mean(totSales), by = c("YEARMONTH", "Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/%100, YEARMONTH %% 100, 1, sep = "-"), "%Y-%m-%d")
][Store_type %in% c("Trial", "Control"), ]
```

Control store 95th percentile
```{r}
pastSales_Controls95 <- pastSales[Store_type == "Control",
][, totSales := totSales * (1 + stdDev * 2)
][, Store_type := "Control 95th % confidence interval"]
## Control store 5th percentile
pastSales_Controls5 <- pastSales[Store_type == "Control",
][, totSales := totSales * (1 - stdDev * 2)
][, Store_type := "Control 5th % confidence interval"]
trialAssessment <- rbind(pastSales, pastSales_Controls95, pastSales_Controls5)
## Plotting these in graph
ggplot(trialAssessment, aes(TransactionMonth, totSales, color = Store_type)) +
geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,],
aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 ,
ymax = Inf, color = NULL), show.legend = FALSE) + 
  geom_line(aes(linetype = Store_type)) + 
  labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")
```

Scale pre-trial control store customers to match pre-trial trial store customers
```{r}
scalingFactorForControlCust <- preTrialMeasures[STORE_NBR == trial_store &
YEARMONTH < 201902, sum(nCustomers)]/preTrialMeasures[STORE_NBR ==
control_store & YEARMONTH < 201902, sum(nCustomers)]

## Apply the scaling factor
measureOverTimeCusts <- measureOverTime
scaledControlCustomers <- measureOverTimeCusts[STORE_NBR == control_store,
][ , controlCustomers := nCustomers * scalingFactorForControlCust
][, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store,"Control", "Other stores"))
]
```

Calculate the absolute percentage difference between scaled control sales and trial sales
```{r}
percentageDiff <- merge(scaledControlCustomers[, c("YEARMONTH","controlCustomers")],measureOverTime[STORE_NBR == trial_store, c("nCustomers", "YEARMONTH")],
by = "YEARMONTH")[, percentageDiff := abs(controlCustomers-nCustomers)/controlCustomers]

## As our null hypothesis is that the trial period is the same as the pre-trial period,

stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff])
degreesOfFreedom <- 7 
## Trial and control store number of customers
pastCustomers <- measureOverTimeCusts[, nCusts := mean(nCustomers), by = c("YEARMONTH", "Store_type")
][Store_type %in% c("Trial", "Control"), ]

```

Control store 95th percentile
```{r}
pastCustomers_Controls95 <- pastCustomers[Store_type == "Control",
][, nCusts := nCusts * (1 + stdDev * 2)
][, Store_type := "Control 95th % confidence interval"]
## Control store 5th percentile
pastCustomers_Controls5 <- pastCustomers[Store_type == "Control",
][, nCusts := nCusts * (1 - stdDev * 2)
][, Store_type := "Control 5th % confidence interval"]
## Combine the tables pastSales, pastSales_Controls95, pastSales_Controls5
trialAssessment <- rbind(pastCustomers, pastCustomers_Controls95,pastCustomers_Controls5)
## Plotting these in one nice graph
ggplot(trialAssessment, aes(TransactionMonth, nCusts, color = Store_type)) +
  geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,],
aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 ,
ymax = Inf, color = NULL), show.legend = FALSE) + geom_line() + 
labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month")
```

## Conclusion

We've found control stores 233, 155, 237 for trial stores 77, 86 and 88 respectively. The results for trial stores 77 and 88 during the trial period show a significant difference in at least two of the three trial months but this is not the case for trial store 86. We can check with the client if the implementation of the trial was different in trial store 86 but overall, the trial shows a significant increase in sales.
